Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2024-09-09 closed_monoidal/closed_monoidal.tex
some updates

\begin{definition}[Definition 6.2.1~\cite{Borceux_1994}]
  Let $\mathcal{V} = (\mathcal{V}_{0}, I_{\mathcal{V}}, \otimes, l, r, a, \sigma)$ be a closed symmetric monoidal category where $l$ and $r$ are left and right unitors and $\sigma$ is a symmetry.
  We define a $\mathcal{V}$-enriched category $\enriched{C}$ consisting of the following data
  \begin{enumerate}
    \item a class of objects $\obj{\enriched{C}}$
    \item for every pair of objects $A,B \in \obj{\enriched{C}}$ --- an (hom-)object $\obj{\enriched{C}}(A,B) \in \mathcal{V}$
    \item for every triple of objects $A,B,C \in \obj{\enriched{C}}$ --- a composition morphism in $\mathcal{V}$
    \[
    c_{A,B,C} : \enriched{C}(A,B) \otimes \enriched{C}(B,C) \to \enriched{C}(A,C)
    \]
    \item for every object $A \in \obj{\enriched{C}}$ --- a unit morphism
    \[
      u_{A} : I_{\mathcal{V}} \to \enriched{C}(A,A)
    \]
  \end{enumerate}
  such that the following diagrams commute (associativity and unitality)
  \[
  \begin{tikzcd}
    {(\enriched{C}(A,B) \otimes \enriched{C}(B,C)) \otimes \enriched{C}(C,D)} \arrow[rr, "{c_{A,B,C} \otimes 1}"] \arrow[d, "{a_{\enriched{C}(A,B),\enriched{C}(B,C),\enriched{C}(C,D)}}"']   &  & {\enriched{C}(A,C) \otimes \enriched{C}(C,D)} \arrow[dd, "{c_{A,C,D}}"]                                                                                     \\
    {\enriched{C}(A,B) \otimes (\enriched{C}(B,C) \otimes \enriched{C}(C,D))} \arrow[d, "{1\otimes c_{B,C,D}}"']                                             &  &                                                                                                                                       \\
    {\enriched{C}(A,B) \otimes \enriched{C}(B,D)} \arrow[rr, "{c_{A,B,D}}"]                                                                       &  & {\enriched{C}(A,D)}                                                                                                                             
  \end{tikzcd}
  \]

@@

@@

\begin{proposition}
  $\catname{SLat}$ is a closed symmetric monoidal category.
\end{proposition}
\begin{proof}
  The tensor product of two semilattices $S_{1}$ and $S_{2}$ is defined as follows.
  $S_{1} \otimes S_{2}$ consists of pairs $(s_1,s_2)$ $s_{1} \in S_{1}$, $s_{2} \in S_{2}$ quotiented by commutativity, idempotence and associativity and additionally by the following relations
  \begin{itemize}
    \item $(s_{1} +_{S_{1}} s_{1}',s_{2}) \cong (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
    \item $(s_{1}, s_{2} +_{S_{2}} s_{2}') \cong (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
  \end{itemize}

@@

\begin{definition}[Isomorphism in $\enriched{C}$]

@@

TODO: I took this definition from \url{https://mathoverflow.net/questions/326575/isomorphisms-in-enriched-categories#:~:text=equals%20id%3A1%E2%86%92,C(X%2CY).} 
Need a proper reference. Riehl, for example, only defines isomorphism of $A$ $B$ in terms of objects of the underlying category.

@@

\begin{proposition}
  The forgetful functor $U : \catname{SLat} \to \catname{Set}$ has a left adjoint free functor $F : \catname{Set} \to \catname{SLat}$.
\end{proposition}
\begin{proof}
  The functor is defined by letting $F(A) = \coprod_{A} I$ where the latter is a coproduct which is a `free' product of semilattices defined as follows.
  The elements of $S_{1} \coprod S_{2}$ are sequences $s_{1} + s_{2} + \ldots + s_{n}$ where each $s_{i}$ is either from $S_{1}$ or $S_{2}$ quotiented by all relations in $S_{1}$ and $S_{2}$.
  In particular, $F(I) = I_{\catname{SLat}}$ and therefore such functor is monoidal with $F(A \otimes B) \cong F(A) \otimes F(B)$.
\end{proof}

TODO: add a note that this follows from the definition of a copower.

\begin{definition}[Proposition 6.4.7~\cite{Borceux_1994}]
  Let $\mathcal{C}$ be a small category.
  Freely enriched $\catname{SLat}$-category $\enriched{C}$ is defined as follows
  \begin{enumerate}
    \item $\obj{\enriched{C}} = \obj{\mathcal{C}}$;
    \item $\enriched{C}(A,B) = F(\mathcal{C}(A,B))$, where $F$ is the left adjoint to $U : \catname{SLat} \to \catname{Set}$;
    \item $c_{A,B,C} : \enriched{C}(A,B) \otimes \enriched{C}(B,C) \to \enriched{C}(A,C)$ is the composite
    \[
    F(\mathcal{C}(A,B)) \otimes F(\mathcal{C}(B,C)) \cong F(\mathcal{C}(A,B) \otimes \mathcal{C}(B,C)) \xrightarrow{F(;)} F(\mathcal{C}(A,C))
    \] 
    where the arrow is the image of the ordinary composition
    \item $u_{A} : I_{\catname{SLat}} \to \enriched{C}(A,A)$ is the composite
    \[
    I_{\catname{SLat}} \cong F(I) \to F(\mathcal{C}(A,A))
    \]
    where the arrow is $* \mapsto 1_{A}$.
  \end{enumerate}
\end{definition}

@@

\begin{example}
  Consider a monoidal bifunctor $\otimes : \mathcal{C} \times \mathcal{C} \to \mathcal{C}$ on $\mathcal{C}$.
  It induces an enriched monoidal bifunctor $\overline{\otimes}$ on $\enriched{C}$.
  \begin{enumerate}
    \item $A \overline{\otimes} B \coloneq A \otimes B$;
    \item $\overline{l} : A \overline{\otimes} I \to A \coloneq I \to \enriched{C}(A \overline{\otimes} I, I)$, i.e. $\overline{l}$ picks out $l$ out of $\enriched{C}(A \overline{\otimes} I, I)$;
    \item similarly for $\overline{r}$, $\overline{\alpha}$.
  \end{enumerate}
  $\otimes_{A,B,C,D} : \mathcal{C}(A,B) \times \mathcal{C}(C,D) \to \mathcal{C}(A \otimes C, B \otimes D)$ becomes $\overline{\otimes}_{A,B,C,D} : \enriched{C}(A,B) \times \enriched{C}(C,D) \to \enriched{C}(A \overline{\otimes} C, B \overline{\otimes} D)$.
  In particular, $f + g \;\overline{\otimes}\; h + k = (f \;\overline{\otimes}\; h + k) + (g \;\overline{\otimes}\; h + k) = (f \;\overline{\otimes}\; k) + (f \;\overline{\otimes}\; h) + (g \;\overline{\otimes}\; h) + (g \;\overline{\otimes}\; k)$, where $+$ binds tighter than $\overline{\otimes}$.
\end{example}

@@

\begin{proposition}
\label{prop:naturality_preserved}
  For a natural transformation $\alpha : T \to S$ for $T,S : \mathcal{C} \to \mathcal{D}$ there is a $\catname{SLat}$-natural transformation $\overline{\alpha} : \overline{T} \to \overline{S}$ where $\overline{T},\overline{S} : \enriched{C} \to \enriched{D}$.
\end{proposition}
\begin{proof}
  The component of a natural transformation $\overline{\alpha}_{A}$ is given by $I_{\catname{SLat}} \to \enriched{D}(\overline{T}A, \overline{S}A)$, i.e. a morphism that picks out the original $\alpha_{A}$.
  We need to check if so defined $\overline{\alpha}$ is $\catname{SLat}$-natural.
  This follows from the following commutativity diagrams
  \[
  \adjustbox{width=\linewidth}{
  \begin{tikzcd}
	&& {\enriched{C}(A,B) \otimes I_{\catname{SLat}}} && {\enriched{D}(\overline{T}A,\overline{T}B) \otimes \enriched{D}(\overline{T}B, \overline{S}B)} &&&&& {f \otimes *} && {\overline{T}f \otimes F\alpha_{B}} \\
	{\enriched{C}(A,B)} &&&&&& {\enriched{D}(\overline{T}A,\overline{S}B)} && f &&&& {F\alpha_{A};\overline{S}f = \overline{T}f ;F\alpha_{B}} \\
	&& {I_{\catname{SLat}} \otimes \enriched{C}(A,B)} && {\enriched{D}(\overline{T}A,\overline{S}A) \otimes \enriched{D}(\overline{S}A,\overline{S}B)} &&&&& {* \otimes f} && {F\alpha_{A} \otimes \overline{S}f}
	\arrow["{\overline{T}_{A,B} \otimes \overline{\alpha_{B}}}", from=1-3, to=1-5]
	\arrow["{c_{\overline{T}A,\overline{T}B,\overline{S}B}}", from=1-5, to=2-7]
	\arrow[from=1-10, to=1-12]
	\arrow[from=1-12, to=2-13]
	\arrow[from=2-1, to=1-3]
	\arrow[from=2-1, to=3-3]
	\arrow[from=2-9, to=1-10]
	\arrow[from=2-9, to=3-10]
	\arrow["{\overline{\alpha_{A}} \otimes \overline{S}(A,B)}", from=3-3, to=3-5]
	\arrow["{c_{\overline{T}A,\overline{S}A,\overline{T}B}}"', from=3-5, to=2-7]
	\arrow[from=3-10, to=3-12]
	\arrow[from=3-12, to=2-13]
\end{tikzcd}}
\]
The diagram on the right is the component-wise rendering of the diagram on the left.
It commutes because it is essentially a lifting of the following diagram (naturality of $\alpha$) using free enrichment functor $F$.

@@

\begin{proposition}

Functor $F: \mathcal{Cat} \to \catname{SLat}-\mathcal{Cat}$ is a 2-functor.
\end{proposition}
\begin{proof}
  This is implied by Propositions~\ref{prop:functoriality_preserved} and \ref{prop:naturality_preserved}.
\end{proof}

The above proposition means, in particular, that an ordinary adjunction gives rise to a corresponding $\catname{SLat}$-enriched adjuntion.
The latter, however, could be also shown Equivalently as below.

\begin{proposition}
  Given an adjunction $\mathcal{C}(LA,B) \cong \mathcal{D}(A,RB)$ of plain functors we also get an $\catname{SLat}$-enriched adjunction $\enriched{C}(\overline{L}A,B) \cong \enriched{D}(A, \overline{R}B)$.
  Because $\overline{L}$ acts on objects as $L$ we will further omit the bar.
\end{proposition}
\begin{proof}
  The first adjunction implies the existence of natural isomorphisms (in $A$ and $B$) between $\mathcal{C}(L-,B)$ and $\mathcal{D}(-,RB)$ and $\mathcal{C}(LA,-)$ and $\mathcal{D}(A,R-)$ respectively.
  Naturality in $B$ is by definition the commutativity of the following diagram
  \begin{equation}
    \label{diag:naturality_plain}
  \begin{tikzcd}
	{\mathcal{C}(LA,B)} && {\mathcal{D}(A,RB)} \\
	\\
	{\mathcal{C}(LA,B')} && {\mathcal{D}(A,RB')}
	\arrow["{\alpha_{B}}"', from=1-1, to=1-3]
	\arrow["{\mathcal{C}(LA,f)}"', from=1-1, to=3-1]
	\arrow["{\mathcal{D}(A,Rf)}", from=1-3, to=3-3]
	\arrow["{\alpha_{B'}}", from=3-1, to=3-3]
  \end{tikzcd}
\end{equation}
  and similarly for the naturality in $A$.
  The claim is that $\enriched{C}(LA,B) \cong \enriched{D}(A,RB)$ is an $\catname{SLat}$-adjunction.
  Recall that $\enriched{C}(LA,B) = F(\mathcal{C}(LA,B))$ and similarly for another hom-set and clearly the isomorphism is preserved: given $\alpha$, an isomorphism from $\mathcal{C}(LA,B) \to \mathcal{D}(A,RB)$, the isomorphism from $\enriched{C}(LA,B) \to \enriched{D}(A,RB)$ is $\enriched{\alpha} = F(\alpha)$.
  We need to check that this $\enriched{\alpha}$ is natural.
  This is equivalent to checking if the following diagram commutes
  \[\begin{tikzcd}
	{\enriched{C}(B,B')} && {[\enriched{C}(LA,B),\enriched{C}(LA,B')]} \\
	\\
	{[\enriched{D}(A,RB),\enriched{D}(A,RB')]} && {[\enriched{C}(A,RB),\enriched{D}(A,RB')]}
	\arrow["{\enriched{C}(LA,-)}", from=1-1, to=1-3]
	\arrow["{\enriched{D}(A,R-)}"', from=1-1, to=3-1]
	\arrow["{[1,\overline{\alpha}_{B'}]}", from=1-3, to=3-3]
	\arrow["{[\overline{\alpha_{B}},1]}"', from=3-1, to=3-3]
\end{tikzcd}\]
which component-wise becomes

@@

\[
\tikzfig{./figures/string_diagrams}
\]

@@

The fact of $+$ being associative and commutative justifies the usage of $\ldots$ in the equations.
\end{section}

TODO: string diagrams for functorial boxes and natural transformations (unit-counit ones).

>>>>

\begin{definition}[Definition 6.2.1~\cite{Borceux_1994}]
  Let $\mathcal{V} = (\mathcal{V}_{0}, I_{\mathcal{V}}, \otimes, l, r, a)$ be a monoidal category where $l$ and $r$ are left and right unitors.
  We define a $\mathcal{V}$-enriched category $\enriched{C}$ consisting of the following data
  \begin{enumerate}
    \item a class of objects $\obj{\enriched{C}}$
    \item for every pair of objects $A,B \in \obj{\enriched{C}}$ --- an (hom-)object $\obj{\enriched{C}}(A,B) \in \mathcal{V}$
    \item for every triple of objects $A,B,C \in \obj{\enriched{C}}$ --- a composition morphism in $\mathcal{V}$
    \[
    c_{A,B,C} : \enriched{C}(A,B) \otimes \enriched{C}(B,C) \to \enriched{C}(A,C)
    \]
    \item for every object $A \in \obj{\enriched{C}}$ --- a unit morphism
    \[
      u_{A} : I_{\mathcal{V}} \to \enriched{C}(A,A)
    \]
  \end{enumerate}
  such that the following diagrams commute (associativity and unitality)
  \[
  \begin{tikzcd}
    {(\enriched{C}(A,B) \otimes \enriched{C}(B,C)) \otimes \enriched{C}(C,D)} \arrow[rr, "{c_{A,B,C} \otimes 1}"] \arrow[d, "{a_{\enriched{C}(A,B),\enriched{C}(B,C),\enriched{C}(C,D)}}"']   &  & {\enriched{C}(A,C) \otimes \enriched{C}(C,D)} \arrow[dd, "{c_{A,C,D}}"]                                                                                     \\
    {\enriched{C}(A,B) \otimes (\enriched{C}(B,C) \otimes \enriched{C}(C,D))} \arrow[d, "{1\otimes c_{B,C,D}}"']                                             &  &                                                                                                                                       \\
    {\enriched{C}(A,B) \otimes \enriched{C}(B,D)} \arrow[rr, "{c_{A,B,D}}"]                                                                       &  & {\enriched{C}(A,D)}                                                                                                                             
  \end{tikzcd}
  \]

@@

\begin{remark}

Requiring that a category $\mathcal{V}$ is monoidal is the most basic requirement for enriched category theory.
To make the notion of enriched category more useful, in particular to be able to talk about a free enrichment, we will further require that $\mathcal{V}$ is a closed symmetric monoidal category.
\end{remark}

@@

\begin{proposition}
  $\catname{SLat}$ is a closed symmetric monoidal category.
\end{proposition}
\begin{proof}
  The tensor product of two semilattices $S_{1}$ and $S_{2}$ is defined as follows.
  $S_{1} \otimes S_{2}$ consists of pairs $(s_1,s_2)$ $s_{1} \in S_{1}$, $s_{2} \in S_{2}$ quotiented by commutativity, idempotence and associativity and additionally by the following relations
  \begin{itemize}
    \item $(s_{1} +_{S_{1}} s_{1}',s_{2}) \equiv (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
    \item $(s_{1}, s_{2} +_{S_{2}} s_{2}') \equiv (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
  \end{itemize}

@@

\begin{definition}[Isomorphism in $\enriched{C}$. Lemma 3.5.12~\cite{Riehl_2014}]

@@

% TODO: I took this definition from \url{https://mathoverflow.net/questions/326575/isomorphisms-in-enriched-categories#:~:text=equals%20id%3A1%E2%86%92,C(X%2CY).} 
% Need a proper reference. Riehl, for example, only defines isomorphism of $A$ $B$ in terms of objects of the underlying category.

@@

\begin{definition}
Forgetful functor $U : \catname{SLat} \to \catname{Set}$ is given by $\catname{SLat}(I_{\catname{SLat}, -}) : \catname{SLat} \to \catname{Set}$.
\end{definition}

Intuitively, the above functor returns the underlying set of a given semilattice $S$ as each morphism from $\{*\} \to S$ picks out an element of $S$.

\begin{proposition}
  The forgetful functor $U : \catname{SLat} \to \catname{Set}$ has a left adjoint free functor $F : \catname{Set} \to \catname{SLat}$.
\end{proposition}
\begin{proof}
  The functor $F$ is defined by letting $F(A) = \coprod_{A} I_{\catname{SLat}}$ where the latter is a coproduct which is a `free' product of semilattices defined as follows.
  The elements of $S_{1} \coprod S_{2}$ are sequences $s_{1} + s_{2} + \ldots + s_{n}$ where each $s_{i}$ is either from $S_{1}$ or $S_{2}$ quotiented by all relations in $S_{1}$ and $S_{2}$.
  The adjunction is then given by the following natural isomorphisms
  \[
  \catname{SLat}(\coprod_{A}(I_{\catname{SLat}}), B) \cong \prod_{A}(\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A,\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A, U(B))
  \]
  The fist isomorphism is given by the fact that hom-functor makes limits into colimits in its first argument, the second being given by the property that $|\catname{Set}(A,B)| = |B|^{|A|} = |\underbrace{B \times \ldots \times B}_{|A|}|$, and the last isomorphism is given by the definition of $U$.
  Furthermore, we have, 
  \[
  F(I) \cong I_{\catname{SLat}}
  \]
  and 
  \begin{align*}
  F(A) \otimes F(B) &\cong (\coprod_{A} I_{\catname{SLat}}) \otimes (\coprod_{B} I_{\catname{SLat}})\\
        &\cong \coprod_{A} (I_{\catname{SLat}} \otimes \coprod_{B} I_{\catname{SLat}})\\
        &\cong \coprod_{A} (\coprod_{B} (I_{\catname{SLat}} \otimes I_{\catname{SLat}}))\\
        &\cong \coprod_{A} (\coprod_{B} I_{\catname{SLat}})\\
        &\cong \coprod_{A \times B} I_{\catname{SLat}}\\
        &\cong F(A \times B)
  \end{align*}
In the above we used the fact that functors $- \otimes X$ and $X \otimes -$ preserve colimits as they are both left-adjoint by symmetry and monoidal closedness of $\catname{SLat}$.
\end{proof}

\begin{remark}
  The above proposition in fact works for any symmetric monoidal closed category $\mathcal{V}$ picked as a base for enrichment (Proposition 6.4.6~\cite{Borceux_1994}) 
\end{remark}

\begin{remark}

  Free semilattice on a set $S$ can be though of as a set of non-empty subsets of $S$.
\end{remark}

\begin{definition}[Proposition 6.4.7~\cite{Borceux_1994}]
\label{def:free_slat}
  Let $\mathcal{C}$ be a small category.
  Freely enriched $\catname{SLat}$-category $\enriched{C}$ is defined as follows
  \begin{enumerate}
    \item $\obj{\enriched{C}} = \obj{\mathcal{C}}$;
    \item $\enriched{C}(A,B) = F(\mathcal{C}(A,B))$, where $F$ is the left adjoint to $U : \catname{SLat} \to \catname{Set}$;
    \item $c_{A,B,C} : \enriched{C}(A,B) \otimes \enriched{C}(B,C) \to \enriched{C}(A,C)$ is the composite $\cong ; F(;)$ below
    \[
    F(\mathcal{C}(A,B)) \otimes F(\mathcal{C}(B,C)) \xrightarrow{\cong} F(\mathcal{C}(A,B) \otimes \mathcal{C}(B,C)) \xrightarrow{F(;)} F(\mathcal{C}(A,C))
    \] 
    where the rightmost arrow is the image of the ordinary composition;
    \item $u_{A} : I_{\catname{SLat}} \to \enriched{C}(A,A)$ is the composite
    \[
    I_{\catname{SLat}} \xrightarrow{\cong} F(I) \to F(\mathcal{C}(A,A))
    \]
    where the rightmost arrow is $* \mapsto 1_{A}$.
  \end{enumerate}
\end{definition}

This is again a specialisation of a more general definition of a free $\mathcal{V}$-category (Proposition 6.4.7~\cite{Borceux_1994}).

@@

\begin{example}
  Let $\mathcal{C} = (\mathcal{C}_{0},I,\otimes, l, r, \alpha)$ be a monoidal category and consider the bifunctor $\otimes : \mathcal{C} \times \mathcal{C} \to \mathcal{C}$ on $\mathcal{C}$.
  It induces an enriched monoidal category $\enriched{C} = (\enriched{C}_{0}, I, \enriched{\otimes}, \overline{l}, \overline{r}, \enriched{\alpha})$ such that
  \begin{enumerate}
    \item $A \overline{\otimes} B \coloneq A \otimes B$;
    \item $\overline{l} : A \overline{\otimes} I \to A \coloneq I \to \enriched{C}(A \overline{\otimes} I, I)$, i.e. $\overline{l}$ picks out $l$ out of $\enriched{C}(A \overline{\otimes} I, I)$;
    \item similarly for $\overline{r}$, $\overline{\alpha}$.
  \end{enumerate}
  $\otimes_{A,B,C,D} : \mathcal{C}(A,B) \times \mathcal{C}(C,D) \to \mathcal{C}(A \otimes C, B \otimes D)$ becomes $\overline{\otimes}_{A,B,C,D} : \enriched{C}(A,B) \times \enriched{C}(C,D) \to \enriched{C}(A \overline{\otimes} C, B \overline{\otimes} D)$.
  In particular, $f + g \;\overline{\otimes}\; h + k = (f \;\overline{\otimes}\; h + k) + (g \;\overline{\otimes}\; h + k) = (f \;\overline{\otimes}\; k) + (f \;\overline{\otimes}\; h) + (g \;\overline{\otimes}\; h) + (g \;\overline{\otimes}\; k)$, where $+$ binds tighter than $\overline{\otimes}$.
\end{example}

@@

\begin{proposition}
\label{prop:naturality_preserved}
  For a natural transformation $\alpha : T \to S$ for $T,S : \mathcal{C} \to \mathcal{D}$ there is a $\catname{SLat}$-natural transformation $\overline{\alpha} : \overline{T} \to \overline{S}$ where $\overline{T},\overline{S} : \enriched{C} \to \enriched{D}$.
\end{proposition}
\begin{proof}
  Recall that an ordinary natural transformation is in fact a $\catname{Set}$-enriched natural transformation and has the form $\alpha_{A} : I \to \mathcal{D}(TA,SA)$.
  We then obtain an $\catname{SLat}$-natural transformation $\overline{\alpha_{A}} = F(\alpha_{A}) : F(I) \to F(\mathcal{D}(TA,SA))$, or $\overline{\alpha_{A}} : I_{\catname{SLat}} \to \enriched{D}(TA,SA)$ given as $* \mapsto F\alpha_{A}$.
  Again, we can use $TA$ or $\overline{T}A$ interchangeably as their action on objects is the same.
  % The component of a natural transformation $\overline{\alpha}_{A}$ is given by $I_{\catname{SLat}} \to \enriched{D}(\overline{T}A, \overline{S}A)$, i.e. a morphism that picks out the original $\alpha_{A}$.
  We need to check if so defined $\overline{\alpha}$ is $\catname{SLat}$-natural.
  This follows from the following commutativity diagrams
  \[
  \adjustbox{width=\linewidth}{
  \begin{tikzcd}
	&& {\enriched{C}(A,B) \otimes I_{\catname{SLat}}} && {\enriched{D}(\overline{T}A,\overline{T}B) \otimes \enriched{D}(\overline{T}B, \overline{S}B)} &&&&& {f \otimes *} && {\overline{T}f \otimes F\alpha_{B}} \\
	{\enriched{C}(A,B)} &&&&&& {\enriched{D}(\overline{T}A,\overline{S}B)} && f &&&& {F\alpha_{A};\overline{S}f = \overline{T}f ;F\alpha_{B}} \\
	&& {I_{\catname{SLat}} \otimes \enriched{C}(A,B)} && {\enriched{D}(\overline{T}A,\overline{S}A) \otimes \enriched{D}(\overline{S}A,\overline{S}B)} &&&&& {* \otimes f} && {F\alpha_{A} \otimes \overline{S}f}
	\arrow["{\overline{T}_{A,B} \otimes \overline{\alpha_{B}}}", from=1-3, to=1-5]
	\arrow["{c_{\overline{T}A,\overline{T}B,\overline{S}B}}", from=1-5, to=2-7]
	\arrow[from=1-10, to=1-12]
	\arrow[from=1-12, to=2-13]
	\arrow[from=2-1, to=1-3]
	\arrow[from=2-1, to=3-3]
	\arrow[from=2-9, to=1-10]
	\arrow[from=2-9, to=3-10]
	\arrow["{\overline{\alpha_{A}} \otimes \overline{S}(A,B)}", from=3-3, to=3-5]
	\arrow["{c_{\overline{T}A,\overline{S}A,\overline{T}B}}"', from=3-5, to=2-7]
	\arrow[from=3-10, to=3-12]
	\arrow[from=3-12, to=2-13]
\end{tikzcd}}
\]
The diagram on the right is the component-wise rendering of the diagram on the left.
It commutes because it is essentially a lifting of the following diagram (naturality of $\alpha$) using free enrichment functor $F$.

@@

\begin{proposition}
Functor $F : \catname{Set} \to \catname{SLat}$ induces a 2-functor $F^{2}: \catname{Cat} \to \catname{SLat}-\catname{Cat}$ from the category of small categories to the category of small categories enriched over a category of semilattices.
\end{proposition}
\begin{proof}
  The functoriality of $F^{2}$ is given by Definition~\ref{def:free_slat} and 2-functoriality is implied by Propositions~\ref{prop:functoriality_preserved} and \ref{prop:naturality_preserved}.
\end{proof}

The above proposition means, in particular, that an ordinary adjunction gives rise to a corresponding $\catname{SLat}$-enriched adjuntion.
The latter, however, could be also shown equivalently as below.

\begin{proposition}
  Given an adjunction $\mathcal{C}(LA,B) \cong \mathcal{D}(A,RB)$ of plain functors $L : \mathcal{D} \to \mathcal{C}$ and $R : \mathcal{C} \to \mathcal{D}$, we also get an $\catname{SLat}$-enriched adjunction $\enriched{C}(\overline{L}A,B) \cong \enriched{D}(A, \overline{R}B)$ of $\catname{SLat}$-functors $\overline{L} : \enriched{D} \to \enriched{C}$ and $\overline{R} : \enriched{C} \to \enriched{D}$.
  Because $\overline{L}$ acts on objects as $L$ we will further omit the bar.
\end{proposition}
\begin{proof}
  The first adjunction implies the existence of natural isomorphisms (in $A$ and $B$) between $\mathcal{C}(L-,B)$ and $\mathcal{D}(-,RB)$ and $\mathcal{C}(LA,-)$ and $\mathcal{D}(A,R-)$ respectively.
  Naturality in $B$ is by definition the commutativity of the following diagram
  \begin{equation}
    \label{diag:naturality_plain}
  \begin{tikzcd}
	{\mathcal{C}(LA,B)} && {\mathcal{D}(A,RB)} \\
	\\
	{\mathcal{C}(LA,B')} && {\mathcal{D}(A,RB')}
	\arrow["{\alpha_{B}}"', from=1-1, to=1-3]
	\arrow["{\mathcal{C}(LA,f)}"', from=1-1, to=3-1]
	\arrow["{\mathcal{D}(A,Rf)}", from=1-3, to=3-3]
	\arrow["{\alpha_{B'}}", from=3-1, to=3-3]
  \end{tikzcd}
\end{equation}
  and similarly for the naturality in $A$.
  The claim is that $\enriched{C}(LA,B) \cong \enriched{D}(A,RB)$ is an $\catname{SLat}$-adjunction.
  Recall that $\enriched{C}(LA,B) = F(\mathcal{C}(LA,B))$ and similarly for another hom-set and clearly the isomorphism is preserved: given $\alpha$, an isomorphism from $\mathcal{C}(LA,B) \to \mathcal{D}(A,RB)$, the isomorphism from $\enriched{C}(LA,B) \to \enriched{D}(A,RB)$ is $\enriched{\alpha} = F(\alpha)$.
  We need to check that this $\enriched{\alpha}$ is natural.
  This is equivalent to checking if the following diagram commutes
  \[\begin{tikzcd}
	{\enriched{C}(B,B')} && {[\enriched{C}(LA,B),\enriched{C}(LA,B')]} \\
	\\
	{[\enriched{D}(A,RB),\enriched{D}(A,RB')]} && {[\enriched{C}(A,RB),\enriched{D}(A,RB')]}
	\arrow["{\enriched{C}(LA,-)}", from=1-1, to=1-3]
	\arrow["{\enriched{D}(A,R-)}"', from=1-1, to=3-1]
	\arrow["{[1,\overline{\alpha}_{B'}]}", from=1-3, to=3-3]
	\arrow["{[\overline{\alpha_{B}},1]}"', from=3-1, to=3-3]
\end{tikzcd}\]
which component-wise becomes

@@

\[
\adjustbox{scale=0.7}{
\tikzfig{./figures/string_diagrams}
}
\]

@@

The fact of $+$ being associative and commutative justifies the usage of $\ldots$ in the equations.

Then we can reuse the diagrams for functorial boxes~\cite{GhicaZanassiStringLambda}.
Given a functor $F : \mathcal{C} \to \mathcal{D}$ and a morphism $f \in C(A,B)$ we can depict the application of $Ff$ as below (left) together with the functoriality equations in Figure~\ref{fig:functoriality_laws}

\[
  \adjustbox{scale=0.7}{
  \tikzfig{./figures/functorial_string}
  }
\]

\begin{figure}
  \begin{subfigure}[c]{0.4\linewidth}
    \[
    \adjustbox{scale=0.7}{
    \tikzfig{./figures/functoriality_eq_1_string}
    }
    \]
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.4\linewidth}
    \[
    \adjustbox{scale=0.7}{
    \tikzfig{./figures/functoriality_eq_2_string}
    }
    \]
  \end{subfigure}
  \caption{String diagrammatic equations for functoriality}
  \label{fig:functoriality_laws}
\end{figure}
\end{section}

A functor $\overline{F}$ induced by the functor $F$ has the same diagrammatic rendering with the addition of an extra equation.

\[
\adjustbox{scale=0.7}{
\tikzfig{./figures/functoriality_enriched_eq}
}
\]

Since a component of a natural transformation $\overline{\alpha}$ between two $\catname{SLat}$-enriched functors $\overline{F} : \enriched{C} \to \enriched{D}$ and $\overline{G}: \enriched{C} \to \enriched{D}$ has the form $\overline{\alpha_{A}} : I_{\catname{SLat}} \to \enriched{D}(FA, GA)$, it can be identified with the element of $\enriched{D}(FA,GA)$, i.e. we can use the syntax for plain morphism to depict components of natural transformations.
This allows us to reuse the whole syntax machinery for Cartesian closed categories for their free $\catname{SLat}$-enriched version which induce an additional equation

\[
\adjustbox{scale=0.7}{
  \tikzfig{./figures/lambda_enriched_eq}
}
\]

We can similarly lift a trace from a traced monoidal category $\mathcal{C}$ to $\enriched{C}$.
Recall that a trace is an operator
\[
\text{tr}^{X} : \mathcal{C}(A \otimes X, B \otimes X) \to \mathcal{C}(A,B)
\]
\textit{i.e.}, a mapping between hom-sets, such that it satisfies the equations
\begin{enumerate}
  \item \textit{Tightening} (naturality in $A,B$) For all $A, B, C, D, X \in \mathcal{C}$, $g : A \to B$, $f : B \otimes X \to C \otimes X$, $h : C \to D$
  \[
  \text{tr}^{X}((h \otimes id_{X});f;(g \otimes id_{X})) = h;\text{tr}^{X}(f);g
  \]
  \item \textit{Sliding} (naturality in $X$) For all $A,B,X,Y \in \mathcal{C}$, $f : A \otimes X \to B \otimes Y$, $g : Y \to X$
  \[
   \text{tr}^{X}(f;(id_{B} \otimes g)) = \text{tr}^{Y}((id_{A} \otimes g); f) 
  \]
  \item \textit{Vanishing} For all $A,B,X,Y \in \mathcal{C}$,
  \[
  \text{tr}^{I}(f) = f,\; \forall f : A \to B \qquad \text{and} \qquad \text{tr}^{X \otimes Y}(f) = \text{tr}^{X}(\text{tr}^{Y}(f)), \; \forall f : A \otimes X \otimes Y \to B \otimes X \otimes Y
  \]
  \item \textit{Strength} For all $A,B,C,D,X \in \mathcal{C}$, $f : C \otimes X \to D \otimes X$, $g : A \to B$
    \[
    \text{tr}^{X}(g \otimes f) = g \otimes \text{tr}^{X}(f)
    \] 
  \item \textit{Yanking} (when $\mathcal{C}$ is symmetric)
    \[
    \text{tr}^{X}(\sigma_{X,X}) = id_{X}
    \]
\end{enumerate}

By freely enriching $\mathcal{C}$ we obtain an enriched trace
\[
\text{tr}^{X} : \enriched{C}(A \overline{\otimes} X, B \overline{\otimes} X) \to \enriched{C}(A,B)
\]
and all the equations are preserved with the addition of a new equation
\[
  \text{tr}^{X}(f + g) = \text{tr}^{X}(f) + \text{tr}^{X}(g)
\]
for $f,g : A \overline{\otimes} X \to B \overline{\otimes} X$
TODO: trace?
